Metadata control document

### 2023-12-01 metadata/metadata_control.Rmd
Saving some cleaned data products for use elsewhere

```{r station locations and CTD summaries}
col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
col_types <- cols(
  station = col_character(),
  cast = col_double(),
  month = col_character(),
  day = col_character(),
  year = col_double(),
  time = col_time(format = ""),
  lat_deg = col_double(),
  lat_min = col_double(),
  lat_dir = col_character(),
  lon_deg = col_double(),
  lon_min = col_double(),
  lon_dir = col_character(),
  max_depth = col_double(),
  bottles = col_double())
toDecDeg <- function(deg, min, dir){
  (deg+min/60)*ifelse(dir%in%c("S", "W"), -1, 1)
}
parseCTD <- function(file_path){
  file_path %>%
    read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}
meso_ctddata <- parseCTD("raw_data/MS_CTDsummary_current.txt")
falkor_ctddata <- parseCTD("raw_data/FK_CTDsummary_current.txt") %>%
  filter(time > "2018-03-23 04:01:51")
complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
```

@@

```{r Core water column parsing (.gof files)}
col_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(col_names, c(head(col_names, 6), tail(col_names, 4)))
gofReader <- function(filename){
  # From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
  # Quality Indicators:
  # Flag  Meaning
  # 1    not quality controled
  # 2    good data
  # 3    suspect (i.e.  questionable) data
  # 4    bad data
  # 5    missing data
  # 9    variable not measured during this cast
  raw_gof <- read_table(filename, skip = 4, col_names = col_names,
                        col_types = cols(.default = col_double()))
  qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
    unite("value", sep = "") %>%
    separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
    mutate(across(everything(), as.numeric))
  raw_gof[,qual_names][qual_df!=2] <- NA
  raw_gof %>% select(-matches("^q\\d+$"))
}
meso_core <- gofReader("raw_data/mesoscope.gof") %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  ))
falkor_core <- gofReader("raw_data/fk180310.gof") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  mutate(cruise="FK")
all_core <- bind_rows(meso_core, falkor_core)
```

>>>>

```{r station locations and CTD summaries}
col_names <- c("station", "cast", "month", "day", "year", "time", 
               "lat_deg", "lat_min", "lat_dir", 
                "lon_deg", "lon_min", "lon_dir", "max_depth", "bottles")
col_types <- cols(
  station = col_character(),
  cast = col_double(),
  month = col_character(),
  day = col_character(),
  year = col_double(),
  time = col_time(format = ""),
  lat_deg = col_double(),
  lat_min = col_double(),
  lat_dir = col_character(),
  lon_deg = col_double(),
  lon_min = col_double(),
  lon_dir = col_character(),
  max_depth = col_double(),
  bottles = col_double())
toDecDeg <- function(deg, min, dir){
  (deg+min/60)*ifelse(dir%in%c("S", "W"), -1, 1)
}
parseCTD <- function(file_path){
  file_path %>%
    read_table(skip = 3, col_names = col_names, col_types=col_types) %>%
    mutate(time=strptime(paste(month, day, year, time), format = "%b %d %Y %X")) %>%
    select(-month, -day, -year) %>%
    mutate(lat=mapply(toDecDeg, lat_deg, lat_min, lat_dir)) %>%
    mutate(lon=mapply(toDecDeg, lon_deg, lon_min, lon_dir)) %>%
    select(-matches("^lon_|^lat_"))
}
meso_ctddata <- parseCTD("raw_data/MS_CTDsummary_current.txt")
falkor_ctddata <- parseCTD("raw_data/FK_CTDsummary_current.txt") %>%
  filter(time > "2018-03-23 04:01:51")
complete_ctddata <- bind_rows(meso_ctddata, falkor_ctddata)
write_csv(complete_ctddata, "complete_ctddata.csv")
```

@@

```{r Core water column parsing (.gof files)}
col_names <- c("station", "cast", "rosette", "lat", "lon", "pressure", "temp",
              "sal", "ctd_oxy", "ctd_chl", "theta", "sigma", "oxy_um", "dic", "alk", "ph",
              "po4_um", "no23_um", "sio4_um", "LLN", "LLP", "PC", "PN", "PP",
              "PSi", "chl_ug", "phaeo_ug", "het_abund", 
              "pro_abund", "syn_abund", "euk_abund",
              "q1", "q2", "q3", "q4")
qual_names <- setdiff(col_names, c(head(col_names, 6), tail(col_names, 4)))
gofReader <- function(filename){
  # From ftp://ftp.soest.hawaii.edu/hot/water/Readme.water.jgofs
  # Quality Indicators:
  # Flag  Meaning
  # 1    not quality controled
  # 2    good data
  # 3    suspect (i.e.  questionable) data
  # 4    bad data
  # 5    missing data
  # 9    variable not measured during this cast
  raw_gof <- read_table(filename, skip = 4, col_names = col_names,
                        col_types = cols(.default = col_double()))
  qual_df <- select(raw_gof, matches("^q\\d+$")) %>%
    unite("value", sep = "") %>%
    separate(sep = "", col = "value", into = c(NA, qual_names)) %>%
    mutate(across(everything(), as.numeric))
  raw_gof[,qual_names][qual_df!=2] <- NA
  raw_gof %>% select(-matches("^q\\d+$"))
}
meso_core <- gofReader("raw_data/mesoscope.gof") %>%
  mutate(cruise=case_when(
    station <= 14 ~ "MT",
    station %in% 15:17 ~ "MC",
    TRUE ~ "Unknown"
  )) %>%
  mutate(station=case_when(
    station <= 15 ~ as.character(station),
    station == 16 ~ "L1",
    station == 17 ~ "L2"
  ))
falkor_core <- gofReader("raw_data/fk180310.gof") %>%
  filter(station>28) %>%
  mutate(station=as.character(station)) %>%
  mutate(cruise="FK")
all_core <- bind_rows(meso_core, falkor_core)
write_csv(all_core, "all_core_gof.csv")
```
